---
title: "Practical Machine Learning - Prediction Assignment"
author: "Nick Cotter"
date: "10 April 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Goal

One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants.

The goal of your project is to predict the manner in which they did the exercise. This is the "classe" variable in the training set. You may use any of the other variables to predict with. You should create a report describing how you built your model, how you used cross validation, what you think the expected out of sample error is, and why you made the choices you did. You will also use your prediction model to predict 20 different test cases.

## Background

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har 
(see the section on the Weight Lifting Exercise Dataset).

## Data

The training data for this project are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv


## Libraries

```{r libraries, message=F, warning=F}
require(knitr)
require(caret)
require(randomForest)
```

# Load Data

```{r loaddata, cache=TRUE}
# load training data
pml_training <- read.csv("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv", na.strings=c("NA","#DIV/0!", ""))
# load testing data - for quiz
pml_testing <- read.csv("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv", na.strings=c("NA","#DIV/0!", ""))
dim(pml_training)
dim(pml_testing)
```

# Pre-processing

## Cleaning The Training Data
```{r cleantrainingdata}
# remove columns with only missing values
pml_training <- pml_training[,colSums(is.na(pml_training)) == 0]
pml_testing <- pml_testing[,colSums(is.na(pml_testing)) == 0]
dim(pml_training)
dim(pml_testing)

# remove irrelevant columns
cols_to_remove <- c("X", "user_name", "raw_timestamp_part_1", "raw_timestamp_part_2", "cvtd_timestamp", "new_window", "num_window")
pml_training <- pml_training[, -which(names(pml_training) %in% cols_to_remove)]
pml_testing <- pml_testing[, -which(names(pml_testing) %in% cols_to_remove)]
dim(pml_training)
dim(pml_testing)
```


# Building The Prediction Model

TODO describe the data - esp. the classe variable


## Partition the training data into training & test
```{r partitiondata}
set.seed(1234)
partitionIndices <- createDataPartition(y = pml_training$classe, p=0.75, list=FALSE)
training_set <- pml_training[partitionIndices, ]
testing_set <- pml_training[-partitionIndices, ]
dim(training_set)
dim(testing_set)
```


## Random Forest

TODO elaborate on why we use random forest - classes of output, sample size, nonlinear etc.

```{r predictionmodel}
# model
model <- randomForest(classe ~ ., data=training_set, method="class")
```

```{r testsetaccuracy}
# prediction
prediction <- predict(model, testing_set, type="class")
confusionMatrix(prediction, testing_set$classe)
```


## Predictions For Test Set

```{r quizpredictions}
pml_testing_prediction <- predict(model, pml_testing, type="class")
pml_testing_prediction
```




## TODO building
## TODO cross validation
## TODO errors
## TODO reasoning behind choices made
## TODO predict 20 different test cases (Quiz)